---
title: "Analysis of eyetracking data with a good fixator"
format: html
editor: visual
---

## Summary, based on the graphs below

Overall this report shows that for at least 7 trials out of 120, the eyetracker is way off, with X off by 90 pixels.

Otherwise, things are reasonable esp. after first 500 ms when only fixation period turns on. But by reasonable, I mean that drift or something means that X and Y was off by up to \~20 pixels.

For the weird 7 trials, there aren't many fixations so I should check how much missing data ("blinks") there was, maybe that helps explain why those trials were weird.

## Do basic checks

```{r}
#| echo: false
library(eyelinkReader)
library(dplyr)
library(ggplot2)
widthPix = 800; heightPix = 600
```

Read in good fixator (someone from the lab who we know was looking at the center) EDF file

```{r}
#| echo: false

goodFixator<-file.path("dataForTestingOfCode","exampleOfGoodFixator","tema.EDF")
EDFexample<- goodFixator

EDFstuff <- eyelinkReader::read_edf(EDFexample,
                                    import_samples = TRUE,
                                    sample_attributes = c('time', 'gx', 'gy'))

specialInstructionsTrials <- c(20,30,31,37,42,45,50,54,60)

```

Special instructions to fixator to look not at fixation trials:

```{r}

print(specialInstructionsTrials)

```

## Graphs

Check mean gX sample and mean gY samples

```{r}
#| echo: false
samples<- EDFstuff$samples
avgSampleRelCtr<- samples %>% filter(! trial %in% specialInstructionsTrials ) %>%
                  summarise(meanX = mean(gxR,na.rm=TRUE), meanY = mean(gyR,na.rm=TRUE))  - 
                             data.frame( meanX=widthPix/2,             meanY= heightPix/2)

if ( any( abs(avgSampleRelCtr) > 40 ) ) { #check if deviation from screen center of average fixation location is greater than 40
  print("Either your screen coordinates are wrong, the eyetracker sucked, or participant didn't look near center much")
}

```

Checked.

Check average FIXATION event location

```{r}
#| echo: false
fixatns <- EDFstuff$fixations
#take the global average which if everything worked right will be near widthPix/2, heightPix/2
avgFix<- fixatns %>% summarise(meanX = mean(gavx), meanY = mean(gavy))  - 
  data.frame( meanX=widthPix/2,     meanY= heightPix/2)
if ( any( abs(avgFix) > 40 ) ) { #check if deviation from screen center of average fixation location is greater than 40 pixels
  print("Either your screen coordinates are wrong, the eyetracker sucked, or participant didn't look near center much")
}
```

Checked

Plot to see how average fixation x changes over time across (non-special) trials

```{r}
#| echo: false
fixatns$distFromFixatn = sqrt( (fixatns$gavx - widthPix/2)^2 + (fixatns$gavy - heightPix/2)^2 )

#Plot to see how it changes over time across trials
fixatns %>% filter(! trial %in% specialInstructionsTrials) %>%
  group_by(trial) %>%
  summarise(gavx = mean(gavx, na.rm = TRUE)) %>%
  ggplot(aes(x = trial, y = gavx)) +
  geom_hline(yintercept = widthPix/2,col="black") +  
  geom_point() #You can see the drift

```

Plot to see how average fixation y changes over time across (non-special) trials

```{r}
#| echo: false

fixatns %>% filter(! trial %in% specialInstructionsTrials) %>%
  group_by(trial) %>%
  summarise(gavy = mean(gavy, na.rm = TRUE)) %>%
  ggplot(aes(x = trial, y = gavy)) +
  geom_hline(yintercept = heightPix/2,col="black") +  
  geom_point()
```

Show drift of average fixation position across trials with 2D plot

```{r}
#| echo: false
fixatns %>% filter(! trial %in% specialInstructionsTrials) %>%
  group_by(trial) %>%
  summarise(gavx = mean(gavx, na.rm = TRUE), gavy = mean(gavy, na.rm = TRUE)) %>%
  ggplot(aes(x = gavx, y = gavy)) +
  geom_hline(yintercept = heightPix/2,col="black") +  
  geom_vline(xintercept = widthPix/2,col="black") +  
  geom_point(aes(color=trial))

#ggplot(fixatns, aes(x=gavx, y=gavy, color=trial) ) +
#  stat_summary(fun="mean",geom="point")

#Implement drift correction,
# by taking average of every trial? Need to decide on limited interval at beginning of trial.
```

I think the first 800 ms at least is fixation period.

Look at x and y traces of trials

```{r}
#| echo: false
fixatns %>% filter(! trial %in% specialInstructionsTrials) %>%
  filter(trial<50) %>% 
  ggplot( aes(x=sttime_rel, y=gavx, color=trial) ) +
  ylab('average x during fixation') + xlab('sttime_rel (ms)') +
  ggtitle('First 50 trials') +
  geom_vline(xintercept = widthPix/2,col="black") +  
  geom_point() + geom_line(aes(group=trial))

fixatns %>% filter(! trial %in% specialInstructionsTrials) %>%
  filter(trial>50) %>% 
  ggplot( aes(x=sttime_rel, y=gavx, color=trial) ) +
  ylab('average x during fixation') + xlab('sttime_rel (ms)') +
  ggtitle('After first 50 trials') +
  geom_vline(xintercept = widthPix/2,col="black") +  
  geom_point() + geom_line(aes(group=trial))

fixatns %>% filter(! trial %in% specialInstructionsTrials) %>%
  filter(trial<50) %>% 
  ggplot( aes(x=sttime_rel, y=gavy, color=trial) ) +
    ggtitle('First 50 trials') +
  ylab('average y during fixation') + xlab('sttime_rel (ms)') +
  geom_vline(xintercept = heightPix/2,col="black") +  
  geom_point() + geom_line(aes(group=trial))

fixatns %>% filter(! trial %in% specialInstructionsTrials) %>%
  filter(trial>50) %>% 
  ggplot( aes(x=sttime_rel, y=gavy, color=trial) ) +
  ggtitle('After first 50 trials') +
  ylab('average y during fixation') + xlab('sttime_rel (ms)') +
  geom_vline(xintercept = heightPix/2,col="black") +  
  geom_point() + geom_line(aes(group=trial))

```

Plot distance from fixation over time

```{r}
#| echo: false

fixatns %>% filter(! trial %in% specialInstructionsTrials) %>%
  ggplot( aes(x=sttime_rel, y=distFromFixatn, color=trial) ) +
  geom_point() + geom_line(aes(group=trial)) +
  ggtitle('Distance from fixation over each trial')
```

The occasional bad stuff happens over the first 100ms, so zoom in Zoom in but show outliers off-scale in red.

```{r}
#| echo: false

minValToShow = 0
maxValToShow = 50
  
fixatns %>% filter(! trial %in% specialInstructionsTrials) %>%
  filter(sttime_rel<1000) %>%
  mutate(outlier = ifelse(distFromFixatn<minValToShow | distFromFixatn>maxValToShow, TRUE, FALSE)) %>%  #determine outliers
  mutate(distFromFixatn =  ifelse(distFromFixatn > maxValToShow, maxValToShow, distFromFixatn)) %>%     #replace outliers
  mutate(distFromFixatn =  ifelse(distFromFixatn < minValToShow, minValToShow, distFromFixatn)) %>%
  ggplot( aes(x=sttime_rel, y=distFromFixatn, color=trial, shape=outlier) ) +
      ylim(minValToShow,maxValToShow) + #restrict axes
      scale_shape_manual(values = c(16, 21)) + #filled circle and unfilled, for outliers
      geom_point(fill='red') + #only the outlier symbol is fillable
      geom_line(aes(group=trial))
```

Zoom in but show outliers off-scale in red.

```{r}
#| echo: false
minValToShow = widthPix/2-100
maxValToShow = widthPix/2+100
fixatns %>% filter(! trial %in% specialInstructionsTrials) %>%
  filter(trial<50) %>% 
  mutate(outlier = ifelse(gavx<minValToShow | gavx>maxValToShow, TRUE, FALSE)) %>%  #determine outliers
  mutate(gavx =  ifelse(gavx > maxValToShow, maxValToShow, gavx)) %>%     #replace outliers
  mutate(gavx =  ifelse(gavx < minValToShow, minValToShow, gavx)) %>%
  ggplot( aes(x=sttime_rel, y=gavx, color=trial, shape=outlier) ) +
  ylab('average x during fixation') + xlab('sttime_rel (ms)') +
  ylim(minValToShow,maxValToShow) + #restrict axes
  scale_shape_manual(values = c(16, 21)) + #filled circle and unfilled, for outliers
  geom_point(fill='red') + #only the outlier symbol is fillable
  geom_hline(yintercept = widthPix/2,col="black") +  
  geom_line(aes(group=trial))
```
